---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

Author: Yong-Han Hank Cheng

This package allows you to generate and compare power spectral density (PSD)
plots given time series data. FFT is used to take a time series data, analyze
the oscillations, and then output the frequencies of these oscillations in the
time series in the form of a PSD plot.

## Installation

```{r, eval = FALSE, warning = FALSE}
# Install the package from GitHub
devtools::install_github("yhhc2/psdr")
```

```{r, warning = FALSE}
# Load package
library("psdr")

```

## Usage

Visit the package's website for function reference: <https://yhhc2.github.io/psdr/>

Visit this vignette for example usage: <https://yhhc2.github.io/psdr/articles/Introduction.html>

## Example

Below is an example of how this package can be used to take a dataframe with
multiple separate time series belonging to 2 categories (A and B), separate out the 
time series, and use the time series to make PSDs and compare the dominant
frequencies between the two categories of signals.

### Load in example dataset

In this example dataset, there are 3 time series for each category. 3 for category
A and 3 for category B. Each time series comes from one session, so there are
6 sessions in total. For each signal, the sampling rate is 100 Hz, which
means a data point is obtained every 0.01 seconds. 

```{r, warning = FALSE}

example_data <- GenerateExampleData()

example_data_displayed <- example_data

colnames(example_data_displayed) <- c("Time in seconds", "Signal", "Session", "Category")

head(example_data_displayed)

```

```{r, warning = FALSE}

#Only works in html, not md. 
rmarkdown::paged_table(example_data_displayed)

```


Here is how the package can be used to take a dataframe containing data from 
all 6 sessions and split it into multiple dataframes, with each dataframe 
containing data from a single session.

```{r}

example_data_windows <- GetHomogeneousWindows(example_data, "Session", c("Session"))

```


### Explore the dataset

Plotting all the time series for category A on a single plot and plotting time 
series data for category B on a single plot shows that the frequencies of 
signals in category A are higher. 

Plot signals for category A
```{r, warning = FALSE}

plot_result <- ggplot2::ggplot(subset(example_data, example_data$Category=="A"), ggplot2::aes(x = Time, y = Signal, colour = Session, group = 1)) + ggplot2::geom_line()

plot_result

```


Plot signals for category B
```{r, warning = FALSE}

plot_result <- ggplot2::ggplot(subset(example_data, example_data$Category=="B"), ggplot2::aes(x = Time, y = Signal, colour = Session, group = 1)) + ggplot2::geom_line()

plot_result


```


This remains true when the time series for each category are averaged.

```{r, warning = FALSE, results = 'hide'}

FirstComboToUse <- list( c(1, 2, 3), c("A") )

SecondComboToUse <- list( c(4, 5, 6), c("B") )

timeseries.results <- AutomatedCompositePlotting(list.of.windows = example_data_windows,
                           name.of.col.containing.time.series = "Signal",
                           x_start = 0,
                           x_end = 999,
                           x_increment = 1,
                           level1.column.name = "Session",
                           level2.column.name = "Category",
                           level.combinations = list(FirstComboToUse, SecondComboToUse),
                           level.combinations.labels = c("A", "B"),
                           plot.title = "Comparing category A and B",
                           plot.xlab = "Time in 0.01 second increments",
                           plot.ylab = "Original units of signal",
                           combination.index.for.envelope = NULL,
                           TimeSeries.PSD.LogPSD = "TimeSeries",
                           sampling_frequency = NULL)

ggplot.obj.timeseries <- timeseries.results[[2]]

ggplot.obj.timeseries

```



### Visualize the frequency contribution of signals

Looking at the time series data, we can tell the frequencies of oscillations are 
different between the time series. To determine which frequencies are contributing
to each time series, we can plot the PSDs for each time series.

PSD for signals in category A
```{r, warning = FALSE}

data1 <- example_data_windows[[1]]
psd_results1 <- MakePowerSpectralDensity(100, data1$Signal)

data2 <- example_data_windows[[2]]
psd_results2 <- MakePowerSpectralDensity(100, data2$Signal)

data3 <- example_data_windows[[3]]
psd_results3 <- MakePowerSpectralDensity(100, data3$Signal)

Frequency <- c(psd_results1[[1]], psd_results2[[1]], psd_results3[[1]])
PSD <- c(psd_results1[[2]], psd_results2[[2]], psd_results3[[2]])
Session <- c(rep(1, length(psd_results1[[1]])), rep(2, length(psd_results1[[1]])), 
             rep(3, length(psd_results1[[1]])))

data_to_plot <- data.frame(Frequency, PSD, Session)

plot_results <- ggplot2::ggplot(data=data_to_plot, ggplot2::aes(x=Frequency, y=PSD, color = factor(Session), group=1)) +
  ggplot2::geom_point() + ggplot2::geom_path() + ggplot2::xlim(0,3)

plot_results

```


PSD for signal in category B
```{r, warning = FALSE}


data1 <- example_data_windows[[4]]
psd_results1 <- MakePowerSpectralDensity(100, data1$Signal)

data2 <- example_data_windows[[5]]
psd_results2 <- MakePowerSpectralDensity(100, data2$Signal)

data3 <- example_data_windows[[6]]
psd_results3 <- MakePowerSpectralDensity(100, data3$Signal)

Frequency <- c(psd_results1[[1]], psd_results2[[1]], psd_results3[[1]])
PSD <- c(psd_results1[[2]], psd_results2[[2]], psd_results3[[2]])
Session <- c(rep(1, length(psd_results1[[1]])), rep(2, length(psd_results1[[1]])), 
             rep(3, length(psd_results1[[1]])))

data_to_plot <- data.frame(Frequency, PSD, Session)

plot_results <- ggplot2::ggplot(data=data_to_plot, ggplot2::aes(x=Frequency, y=PSD, color = factor(Session), group=1)) +
  ggplot2::geom_point() + ggplot2::geom_path() + ggplot2::xlim(0,3)

plot_results

```


To get a single composite PSD for each category, we can take the average.

```{r, warning = FALSE, results = 'hide'}

FirstComboToUse <- list( c(1, 2, 3), c("A") )

SecondComboToUse <- list( c(4, 5, 6), c("B") )

PSD.results <- AutomatedCompositePlotting(list.of.windows = example_data_windows,
                           name.of.col.containing.time.series = "Signal",
                           x_start = 0,
                           x_end = 5,
                           x_increment = 0.01,
                           level1.column.name = "Session",
                           level2.column.name = "Category",
                           level.combinations = list(FirstComboToUse, SecondComboToUse),
                           level.combinations.labels = c("A", "B"),
                           plot.title = "Comparing category A and B",
                           plot.xlab = "Hz",
                           plot.ylab = "(Original units)^2/Hz",
                           combination.index.for.envelope = NULL,
                           TimeSeries.PSD.LogPSD = "PSD",
                           sampling_frequency = 100)

ggplot.obj.PSD <- PSD.results[[2]]

ggplot.obj.PSD

```

If we want to see how the average compares to the individual signals that
make up the average, then we can include an error envelope.

Here is the error envelope added to the category A composite curve.
```{r, warning = FALSE, results = 'hide'}
PSD.results <- AutomatedCompositePlotting(list.of.windows = example_data_windows,
                           name.of.col.containing.time.series = "Signal",
                           x_start = 0,
                           x_end = 5,
                           x_increment = 0.01,
                           level1.column.name = "Session",
                           level2.column.name = "Category",
                           level.combinations = list(FirstComboToUse, SecondComboToUse),
                           level.combinations.labels = c("A", "B"),
                           plot.title = "Comparing category A and B",
                           plot.xlab = "Hz",
                           plot.ylab = "(Original units)^2/Hz",
                           combination.index.for.envelope = 1,
                           TimeSeries.PSD.LogPSD = "PSD",
                           sampling_frequency = 100
                           )

ggplot.obj.PSD <- PSD.results[[2]]

ggplot.obj.PSD
```


Here is the error envelope added to the category B composite curve.
```{r, warning = FALSE, results = 'hide'}
PSD.results <- AutomatedCompositePlotting(list.of.windows = example_data_windows,
                           name.of.col.containing.time.series = "Signal",
                           x_start = 0,
                           x_end = 5,
                           x_increment = 0.01,
                           level1.column.name = "Session",
                           level2.column.name = "Category",
                           level.combinations = list(FirstComboToUse, SecondComboToUse),
                           level.combinations.labels = c("A", "B"),
                           plot.title = "Comparing category A and B",
                           plot.xlab = "Hz",
                           plot.ylab = "(Original units)^2/Hz",
                           combination.index.for.envelope = 2,
                           TimeSeries.PSD.LogPSD = "PSD",
                           sampling_frequency = 100
                           )

ggplot.obj.PSD <- PSD.results[[2]]

ggplot.obj.PSD
```


When the signals are very noisy, it is often times helpful
to log transform the PSD plots. For the example data, this is
not necessary because the signals are very clear. Since amplitudes
are small, log transform are also not helpful here. 
```{r, warning = FALSE, results = 'hide'}
LogPSD.results <- AutomatedCompositePlotting(list.of.windows = example_data_windows,
                           name.of.col.containing.time.series = "Signal",
                           x_start = 0,
                           x_end = 5,
                           x_increment = 0.01,
                           level1.column.name = "Session",
                           level2.column.name = "Category",
                           level.combinations = list(FirstComboToUse, SecondComboToUse),
                           level.combinations.labels = c("A", "B"),
                           plot.title = "Comparing category A and B",
                           plot.xlab = "Hz",
                           plot.ylab = "Log((Original units)^2/Hz)",
                           combination.index.for.envelope = NULL,
                           TimeSeries.PSD.LogPSD = "LogPSD",
                           sampling_frequency = 100
                           )

ggplot.obj.LogPSD <- LogPSD.results[[2]]

ggplot.obj.LogPSD
```



### Comparing frequency contribution of each category

We know there are differences in frequencies of signals between category A and
B, but we want to statistically test if the difference is significant. 

```{r, warning = FALSE}

comparison_results <- PSD.results[[3]]

dominant_freq_for_comparison <- comparison_results[[1]]

kruskal_wallis_test_results <- comparison_results[[2]]

wilcoxon_rank_sum_test_results <- comparison_results[[3]]

```

Since multiple signals are present in each category, we want to 
see if the dominant frequencies in signals of category A are significantly
different from the dominant frequencies in signals of category B

```{r, warning = FALSE}
dominant_freq_for_comparison
```

The comparison can be performed using the Kruskal-Wallis rank sum test. Here,
the p-value indicates the difference is statistically significant. 

```{r, warning = FALSE}
kruskal_wallis_test_results
```

In this example, only two categories are used. However, Kruskal-Wallis rank
sum test can be used for multiple categories, similar to ANOVA. If more than
two categories are used, pair-wise testing using Wilcoxon rank sum exact test
can be used to see which two categories are significantly different.

```{r, warning = FALSE}
wilcoxon_rank_sum_test_results
```


